FitBit Analysis

### 2022-11-24 02_paper/02_study/02_fitbit paper/fitbit_paper.Rmd
update fitbit paper

keywords          : "heart rate; photoplethysmography; wearable electronic device; expertise differences"
wordcount         : "X"

@@

```{r include=FALSE}
# install.packages("needs")

# suppress "summarize" info. 
# if this line is ommitted, each table using the summarize function will be accompanied with a warning from the console
options(dplyr.summarise.inform = FALSE)

library(needs)

needs(tidyverse,
      lubridate,
      janitor,
      viridis,
      grid,
      gridExtra,
      cowplot,
      readxl,
      ARTofR,
      moments,
      ltm,
      sjPlot,
      kableExtra,
      xtable,
      DescTools,
      formattable,
      ggthemes,
      imputeTS,
      rstatix,
      ggpubr)

@@

## Aim of the study
Against this background, we investigated whether heart rate measurements using fitness trackers in a controlled teaching-learning setting are suitable to map differences between (1) anticipation, arousal, and rest phases and (2) between experienced and inexperienced teachers. We expected heart rates to be higher in the arousal phase than in the rest phases and exploratively examine whether experienced teachers have lower heart rates and more steps compared to inexperienced teachers.

@@

# DEMOGRAPHIC DATA
# read in data and select and rename relevant columns with transmute 
demo.data.raw <- 
  read_excel("./data/data_empschul_labor_lehrperson.xlsx") %>% 
  transmute(ID = LI06_05,
            Gender = LI02_01_1,
            Age = LI03_01,
            `Teaching Experience` = LI04_01,
            School_type = LI18,
            Grade = LI16_05,
            Subject = LI10_05,
            Semester = LI11_01,
            Internship = LI13_01,
            `Extracurricular Teaching Experience` = LI14,
            `Secondary Teaching Activities` = LI17
            )

# exclude cases with no fitbit data
demo.data <- demo.data.raw %>% filter(!ID %in% c("126",
                                                 "132"
                                                 )
                                      )

# define expert and novice with ifelse function
demo.data$ID = ifelse(demo.data$ID < 200, "Novice","Expert")

# changing gender to count female gender (male = 2 --> 0)
demo.data$Gender = ifelse(demo.data$Gender == 2, "1","0")

# changing character Gender into numeric
demo.data$Gender <- as.numeric(demo.data$Gender)

# creating new column with total participants to calculate percent of gender
demo.data <- demo.data %>%
  group_by(ID) %>%
  mutate(Sum_participants = n())

# replacing 2 to 1 (Oberschule = Gymnasisium --> Secondary school)
demo.data$School_type[demo.data$School_type == 2] <- "1"

# creating new column to calculate percent of school type
demo.data <- 
  demo.data %>%
  group_by(School_type, ID) %>%
  mutate(Count_Schooltype = sum(n())) %>% 
  mutate(Percent_schooltype = round((Count_Schooltype/Sum_participants) * 100, 
         digits = 2))

# changing character Percent_schooltype into numeric
demo.data$Percent_schooltype <- as.numeric(demo.data$Percent_schooltype)

# changing character Semester into numeric
demo.data$Semester <- as.numeric(demo.data$Semester)

# changing Type of School into character
# LI18:
# 1: Grundschule
# 2: Oberschule
# 3: Gymnasium
# 4: Förderschule/Sondershausen
# 5: Berufsbildende Schule
# -1: andere Schulform
demo.data$School_type <- factor(x = demo.data$School_type,
                   levels = c("1","3","4","5","-1"), # Mögliche Werte # andere Möglichkeit: unique(demo.data$School_type)
                   labels = c("Primary School","Secondary School", "Special Education","Vocational School", "Other Schooltype")) # Bezeichnungen

# changing `Extracurricular Teaching Experience` to count experience (no experience = 2 --> 0)
demo.data$`Extracurricular Teaching Experience` = ifelse(demo.data$`Extracurricular Teaching Experience` == 1, "1","0")

# changing character Extra. Exp. into numeric
demo.data$`Extracurricular Teaching Experience` <- as.numeric(demo.data$`Extracurricular Teaching Experience`)

# changing `Secondary Teaching Activities` to count experience (no experience = 2 --> 0)
demo.data$`Secondary Teaching Activities` = ifelse(demo.data$`Secondary Teaching Activities` == 1, "1","0")

# changing character Extra. Exp. into numeric
demo.data$`Secondary Teaching Activities` <- as.numeric(demo.data$`Secondary Teaching Activities`)

# create new columns that give proportion of Gender
# percent = female gender divided by all female Gender
# demo.data <- demo.data  %>% 
#   group_by(Group) %>% 
#   mutate(Gender_percent = Gender/Sum_participants) %>% 
#   mutate(Gender_percent = round(sum(Gender_percent)*100, digits=0))

# create a basic table (tibble) using tidyverse functions
demo.table <- demo.data %>%
  group_by(ID) %>%
  summarise(N=n(),
            "Gender female in percent" = round(sum(Gender/Sum_participants) * 100, 
                                               digits = 2),
            "M Age in years" = round(mean(Age), 
                                     digits = 2),
            "SD Age in years" = round(sd(Age), 
                                      digits = 2),
            "Min Age in years" = min(Age),
            "Max Age in years" = max(Age)
            )

@@

The sample consisted of *N* = `r length(demo.data$ID)` participants with *n* = `r length(demo.data$ID[demo.data$ID == "Expert"])` expert teachers and *n* = `r length(demo.data$ID[demo.data$ID == "Novice"])` novice teachers.

>>>>

keywords          : "heart rate; photoplethysmography; wearable electronic device; teaching"
wordcount         : "X"

@@

```{r data wrangling, include=FALSE}

library(needs)
needs(ltm,
      broom,
      ppcor,
      jtools,
      lm.beta,
      tidyverse,
      janitor,
      lubridate,
      readxl,
      ggthemes,
      gridExtra,
      imputeTS,
      DescTools,
      cowplot,
      rstatix,
      ggpubr,
      lme4,
      viridis)

# disruption & confidence rating
df_rating <- 
  excel_sheets("data/Coding_SRI.xlsx") %>% 
  map_df(~read_xlsx("data/Coding_SRI.xlsx",.)) %>% # read in data with two sheets
  select(ID, # select relevant columns 
         disruption_factor, # -99 = subject did not notice event; -100 = experimenter did not asked for event
         confident_factor,
         event
         ) %>% 
  mutate(disruption_factor = ifelse(disruption_factor < 0,
                                    yes = NA,
                                    no = disruption_factor
                                    ),
         confident_factor = ifelse(confident_factor < 0,
                                   yes = NA,
                                   no = confident_factor),
         event = as_factor(event)
         )

# demographic data
df_demo <- 
  read_excel("./data/data_empschul_labor_lehrperson.xlsx") %>% # read in excel 
  transmute(ID = LI06_05, # select and rename relevant columns 
            gender = factor(LI02_01_1,
                            levels = 1:2,
                            labels = c("male","female")
                            ),
            age = LI03_01, # 1 = male; 2 = female
            teaching_experience = LI04_01,
            school_type = LI18,
            term = LI11_01, 
            extracurricular_teaching_experience = LI14,
            secondary_teaching_activities = LI17)

# merge two data frames by ID 
df_merge <- merge(df_demo,
                  df_rating,
                  by = "ID") %>% 
  filter(!ID %in% c("126", # exclude cases with no fitbit data (to check see data Heart Rate)
                    "132")
         )

# rm(list = c("df_demo","df_rating"))

data_path <- "./data/heart_rate_data"

main <-
map(.x = dir(path = data_path,
             pattern = ".csv"),
    ~ read_csv(file.path(data_path, .), 
               id = "id",
               col_types = c("t","n")
               ) %>% 
  mutate(time = .$Time - min(.$Time),
         time = as.numeric(time),
         heart_rate = `Heart Rate`,
         ID = id,
         time_span = case_when(str_detect(string = id, pattern = "_a") ~ "after",
           str_detect(string = id, pattern = "_m") ~ "main",
           str_detect(string = id, pattern = "_p") ~ "pre",
           str_detect(string = id, pattern = "_s") ~ "subsequent",
           str_detect(string = id, pattern = "_i") ~ "interview",
           TRUE ~ "overall"
           ),
         ID = str_extract(string = ID,
                          pattern = "[:digit:]{3}"),
         ID = as.numeric(ID)
      ) %>%
      filter(time <= 600) %>%
    select(!c("Time","Heart Rate","id"))
    ) %>%
  bind_rows()

main <-
  left_join(main, df_merge,
            by = "ID")
```

@@

## Aim of the study
Against this background, we investigated whether heart rate measurements using fitness trackers in a controlled teaching-learning setting are suitable to map differences between (1) anticipation, arousal, and rest phases. We expected heart rates to be higher in the arousal phase than in the rest phases.

@@

# define expert and novice with ifelse function
df_demo$ID = ifelse(df_demo$ID < 200, "Novice","Expert")

# changing gender to count female gender (male = 2 --> 0)
df_demo$gender = ifelse(df_demo$gender == 2, "1","0")

# changing character Gender into numeric
df_demo$gender <- as.numeric(df_demo$gender)

# creating new column with total participants to calculate percent of gender
df_demo <- df_demo %>%
  group_by(ID) %>%
  mutate(sum_participants = n())

# replacing 2 to 1 (Oberschule = Gymnasisium --> Secondary school)
df_demo$school_type[df_demo$school_type == 2] <- "1"

# creating new column to calculate percent of school type
df_demo <- 
  df_demo %>%
  group_by(school_type, ID) %>%
  mutate(count_schooltype = sum(n())) %>% 
  mutate(percent_schooltype = round((count_schooltype/sum_participants) * 100, 
         digits = 2))

# changing character percent_schooltype into numeric
df_demo$percent_schooltype <- as.numeric(df_demo$percent_schooltype)

# changing Type of School into character
# LI18:
# 1: Grundschule
# 2: Oberschule
# 3: Gymnasium
# 4: Förderschule/Sondershausen
# 5: Berufsbildende Schule
# -1: andere Schulform
df_demo$school_type <- factor(x = df_demo$school_type,
                   levels = c("1","3","4","5","-1"), # Mögliche Werte # andere Möglichkeit: unique(demo.data$School_type)
                   labels = c("Primary School","Secondary School", "Special Education","Vocational School", "Other Schooltype")) # Bezeichnungen

# changing character semester into numeric
df_demo$term <- as.numeric(df_demo$term)

# changing `extracurricular_teaching_experience to count experience (no experience = 2 --> 0)
df_demo$extracurricular_teaching_experience = ifelse(demo.data$extracurricular_teaching_experience == 1, "1","0")

# changing character extracurricular_teaching_experience into numeric
df_demo$extracurricular_teaching_experience <- as.numeric(df_demo$extracurricular_teaching_experience)

# changing secondary_teaching_activities to count experience (no experience = 2 --> 0)
df_demo$secondary_teaching_activities = ifelse(df_demo$secondary_teaching_activities == 1, "1","0")

# changing character Extra. Exp. into numeric
df_demo$secondary_teaching_activities <- as.numeric(df_demo$secondary_teaching_activities)

# create a basic table (tibble) using tidyverse functions
df_demo <- df_demo %>%
  group_by(ID) %>%
  summarise(N=n(),
            "Gender female in percent" = round(sum(Gender/Sum_participants) * 100, 
                                               digits = 2),
            "M Age in years" = round(mean(Age), 
                                     digits = 2),
            "SD Age in years" = round(sd(Age), 
                                      digits = 2),
            "Min Age in years" = min(Age),
            "Max Age in years" = max(Age)
            )

@@

The sample consisted of *N* = `r length(df_demo$ID)` participants with *n* = `r length(df_demo$ID[df_demo$ID == "Expert"])` expert teachers and *n* = `r length(demo.data$ID[demo.data$ID == "Novice"])` novice teachers.
